An Experimental Evaluation of Hybrid Querying on Vectors (EA\&B)

### 2025-05-01 VLDB2026/main.tex
Update main.tex

	%
	% The abstract is a short summary of the work to be presented in the
	% article.
	\begin{abstract}
		Recent studies demonstrate the significant practical value of hybrid queries, which integrate vector search with structured filters (e.g., attribute and range filtering) for refined retrieval.
		However, current evaluations lack unified benchmarking standards and systematic assessment methodologies. Existing studies not only fail to cover mainstream algorithms but also omit systematic comparisons or in-depth analysis on different methods. To address this issue, we design a complete evaluation framework for hybrid queries. Our study introduces 15 hybrid query algorithms and systematically classifies them based on multiple dimensions, such as index organization and filtering strategy, providing a reference for the categorization of hybrid queries. In experiments, for attribute filtering, we construct standard attribute sets, enabling a unified comparison of algorithms in terms of index construction efficiency, query performance, and robustness. For range filtering, we systematically evaluate performance across the 3 metrics through controlled variation of query ranges. Additionally, we conduct an in-depth analysis of the experimental results based on the underlying principles of the algorithms. The extensive experimental results reveal the strengths and weaknesses of each algorithm. Based on these findings, we develop a set of practical guidelines for algorithm selection, offering reliable references for different application scenarios. Furthermore, we identify potential directions for improvement to address the current limitations of these algorithms.

@@

	Nearest Neighbor (NN) search \cite{cover1967nearest} aims to find the closest vector in a given space and serves as a fundamental algorithm in vector retrieval, with widespread applications in recommendation systems  \cite{monolith}  and image retrieval \cite{wang2012scalable}.
	% 需要添加引用，下面这段
	However, with the exponential growth of data volume and the increasing dimensionality of vectors \cite{weber1998quantitative}, traditional NN search methods struggle to meet the demands of real-time search. To address this issue, researchers turn to more efficient approaches - Approximate NN search ~\cite{arya1998optimal,beis1997shape,gionis1999similarity,malkov2018efficient}. ANN algorithms significantly improve search efficiency by constructing effective indexing structures, albeit at the cost of reduced accuracy.

	Nevertheless, as application scenarios grow increasingly complex, simple ANN search can no longer satisfy all practical needs. For instance, Figure ~\ref{fig:hybrid ANNS} illustrates a case where users on e-commerce platforms search for clothing items by retrieving visually similar products based on an image. Additionally, users may impose further requirements such as price, color, or brand preferences. Such scenarios necessitate a retrieval system capable of simultaneously addressing vector similarity (e.g., product image) and attribute constraints (e.g., brand name) \cite{tian2023approximate}. When the constraint involves a specific attribute value, this problem refers to Attribute Filtering Approximate Nearest Neighbor (AF-ANN) search ~\cite{NHQ,Filtered-diskann} . For example, a user may seek a green piece of clothing. If the constraint involves a range condition, the problem is known as Range Filtering Approximate Nearest Neighbor (RF-ANN) search ~\cite{serf,iRangeGraph}. For instance, filtering clothes priced between 100 and 200. To meet these application demands, hybrid query \cite{JD-e-commerce, analyticdb} techniques emerge. Hybrid queries integrate vector retrieval with conditional filtering, optimizing their interaction to significantly enhance efficiency and flexibility in complex query scenarios.

@@

	\textit{Performance of Algorithms under Multi-Attribute Building (M vs. M).}  
	As shown in Figure \ref{fig:exp_2_1}, IVF-based methods (Puck and CAPS) significantly outperform graph-based methods. The reason for Puck's superior performance is that it filters out irrelevant results during the retrieval process rather than searching for more candidates. CAPS performs as well as Puck. In contrast, FilteredVamana performs the worst because its multi-attribute indexing forces neighbor selection to cover multiple attributes, reducing the efficiency of single-attribute queries. This results in excessive traversal of irrelevant nodes, making it slower than other methods.

@@

	Lastly, our findings show that the distribution and the number of attributes, as well as dataset properties (e.g., dimensionality and LID), significantly impact the performance of attribute filtering methods. Enhancing the robustness of these algorithms under diverse data conditions is another key challenge that warrants further investigation.

	\section{CONCLUSION}
	In this paper, we systematically evaluate hybrid query methods, including various algorithms, vector databases, and libraries. We enrich the dataset with attribute information and design diverse experimental settings. We analyze the performance of these algorithms in depth. We find that these algorithms have different strengths and weaknesses under large-scale data, complex queries, and resource-constrained scenarios. Attribute filtering algorithms show clear limitations in flexibility and Boolean logic support. Range filtering algorithms show weaknesses in multi-attribute queries and high-recall performance. Sometimes, baseline methods like Faiss perform better in high-recall settings. In the future, we need to improve the flexibility and robustness of attribute filtering algorithms and enhance the performance of range filtering algorithms.
%	\textcolor{blue}{
%		In this paper, we evaluate hybrid query methods, including algorithms, vector databases, and libraries. We design experiments to assess the performance of attribute filtering algorithms. We enrich existing datasets with attribute values and create various experimental scenarios to ensure fair comparisons. We run experiments on 7 real-world datasets to analyze the effectiveness of attribute filtering algorithms. We also compare range filtering algorithms using 4 range query datasets. Finally, we analyze the results, summarize key findings, and identify areas for improvement. Our study validates previous research and provides insights for future work.
%	}
%

>>>>

	%
	% The abstract is a short summary of the work to be presented in the
	% article.
	\begin{abstract}
		Recent studies demonstrate the significant practical value of hybrid queries, which integrate vector search with structured filters (e.g., attribute and range filtering) for refined retrieval.
		However, current evaluations lack unified benchmarking standards and systematic assessment methodologies. Existing studies not only fail to cover mainstream algorithms but also omit systematic comparisons or in-depth analysis on different methods. To address this issue, we design a complete evaluation framework for hybrid queries. Our study introduces 15 hybrid query algorithms and systematically classifies them based on multiple dimensions, such as index organization and filtering strategy, providing a reference for the categorization of hybrid queries. In experiments, for attribute filtering, we construct standard attribute sets, enabling a unified comparison of algorithms in terms of index construction efficiency, query performance, and robustness. For range filtering, we also evaluate the algorithm performance across the 3 metrics through controlled variation of query ranges. Additionally, we conduct an in-depth analysis of the experimental results based on the underlying principles of algorithms. Extensive experimental results reveal the strengths and weaknesses of each algorithm. Based on the findings, we develop a set of practical guidelines for algorithm selection, offering reliable references for different application scenarios. Furthermore, we identify potential directions for improvement to address the current limitations of these algorithms.

@@

	Nearest Neighbor (NN) search \cite{cover1967nearest} aims to find the closest vector in a given space and serves as a fundamental algorithm in vector retrieval, with widespread applications in recommendation systems  \cite{monolith}  and image retrieval \cite{wang2012scalable}.
	% 需要添加引用，下面这段
	However, with the exponential growth of data volume and the increasing dimensionality of vectors \cite{weber1998quantitative}, traditional NN search methods struggle to meet the demands of real-time search. To address this issue, studies turn to more efficient approaches - Approximate NN (ANN) search ~\cite{arya1998optimal,beis1997shape,gionis1999similarity,malkov2018efficient}, which significantly improve search efficiency by constructing effective indexing structures, albeit at the cost of reduced accuracy.

	Nevertheless, as application scenarios grow increasingly complex, simple ANN search can no longer satisfy all practical needs. For instance, Fig.~\ref{fig:hybrid ANNS} illustrates a case where users on e-commerce platforms search for clothing items by retrieving visually similar products based on an image. Additionally, users may impose further requirements such as price, color, or brand preferences. Such scenarios necessitate a retrieval system capable of simultaneously addressing vector similarity (e.g., product image) and attribute constraints (e.g., brand name) \cite{tian2023approximate}. When the constraint involves a specific attribute value, this problem refers to Attribute Filtering Approximate Nearest Neighbor (AF-ANN) search ~\cite{NHQ,Filtered-diskann} . For example, a user may seek a green piece of clothing. If the constraint involves a range condition, the problem is known as Range Filtering Approximate Nearest Neighbor (RF-ANN) search ~\cite{serf,iRangeGraph}. For instance, filtering clothes priced between 100 and 200. To meet these application demands, hybrid querying \cite{JD-e-commerce, analyticdb} techniques emerge. Hybrid queries integrate vector retrieval with conditional filtering, optimizing their interaction to significantly enhance efficiency and flexibility in complex query scenarios.

@@

	\textit{Performance of Algorithms under Multi-Attribute Building (M vs. M).}  
	As shown in Figure \ref{fig:exp_2_1}, IVF-based methods (Puck and CAPS) significantly outperform graph-based methods. The reason for Puck's superior performance is that it filters out irrelevant results during the retrieval process rather than searching for more candidates. \textcolor{red}{CAPS performs as well as Puck, but differs by further grouping results by attributes within clusters.} In contrast, FilteredVamana performs the worst because its multi-attribute indexing forces neighbor selection to cover multiple attributes, reducing the efficiency of single-attribute queries. This results in excessive traversal of irrelevant nodes, making it slower than other methods.

@@

	Lastly, our findings show that the distribution and the \textcolor{blue}{ number} of attributes, as well as dataset properties (e.g., dimensionality and LID), significantly impact the performance of attribute filtering methods. Enhancing the robustness of these algorithms under diverse data conditions is another key challenge that warrants further investigation.

	\section{CONCLUSION}
	\textcolor{blue}{In this paper, we systematically evaluate hybrid query methods, including various algorithms, vector databases, and libraries. We enrich the dataset with attribute information and design diverse experimental settings. We analyze the performance of these algorithms in depth. We find that these algorithms have different strengths and weaknesses under large-scale data, complex queries, and resource-constrained scenarios. Attribute filtering algorithms show clear limitations in flexibility and Boolean logic support. Range filtering algorithms show weaknesses in multi-attribute queries and high-recall performance. Sometimes, baseline methods like Faiss perform better in high-recall settings. In the future, we need to improve the flexibility and robustness of attribute filtering algorithms and enhance the performance of range filtering algorithms.}
%	\textcolor{blue}{
%		In this paper, we evaluate hybrid query methods, including algorithms, vector databases, and libraries. We design experiments to assess the performance of attribute filtering algorithms. We enrich existing datasets with attribute values and create various experimental scenarios to ensure fair comparisons. We run experiments on 7 real-world datasets to analyze the effectiveness of attribute filtering algorithms. We also compare range filtering algorithms using 4 range query datasets. Finally, we analyze the results, summarize key findings, and identify areas for improvement. Our study validates previous research and provides insights for future work.
%	}
%
